An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-16 poly-embed-journal/poly-embed-journal.tex
Phase 1 of Lemma X

\section{``Book chapter'' approach}
\subsection*{Definitions and preliminaries}
\begin{definition}[Positive halfspace]
	For any $v\in\reals^\Y_+$, let $H_{v}^+ := \{ (x,c) \in \reals^\Y \times \reals \mid \inprod{x}{v} - c \geq 0\}$ be the \emph{linear positive halfspace}, and $H_{(w,b)}^+ := \{ x \in \reals^{\Y+1} \mid \inprod{x}{w} - b \geq 0\}$ be the \emph{affine positive halfspace}
\end{definition}
Abusing notation, we sometimes let $H_y^+ := \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\begin{definition}[Hyperplane]
	For any $v\in\reals^\Y_+$, let $H_v := \{ (x,c) \in \reals^\Y \times \reals \mid \inprod{x}{v} - c = 0\}$ be the hyperplane generated by $v$.
\end{definition} 
\begin{definition}[$\H$-polyhedron\jessie{Ziegler page 28}]
	An \emph{$\H$-polyhedron} $P$ describes an intersection of closed halfspaces presented in the form
	\begin{equation*}
	P = P(A,z) = \{x \in \reals^d \mid Ax \leq z\} \qquad \textrm{ for some } A \in \reals^{m \times d}, z \in \reals^m~.~
	\end{equation*}
\end{definition}
\begin{definition}[Face, facet]
	Let $P \subseteq \reals^d$ be a convex polytope.
	A linear inequality $\inprod{v}{x} \geq c$ is \emph{valid} for $P$ if it is satisfied for all $x \in P$.
	A \emph{face} of the polytope $P$ is any set of the form
	\begin{equation*}
	F = P \cap \{x \in \reals^d \mid \inprod{v}{x} = c\}~.~
	\end{equation*}
	The dimension of a face is the dimension of its affine hull $\dim(F) := \dim(\affhull(F))$.
	A $(\dim(P) - 1)$-dimensional face is called a facet.
\end{definition}

@@

Set up notation and define $g,G$
\begin{itemize}\setlength{\itemsep}{0pt}
\item Let $f:\simplex\to\reals_+$ be a polyhedral concave function.
\begin{claim}
	We may write $f(p) = \min_{v\in\hat\V} \inprod{p}{v}$ for some finite set $\hat\V \subseteq \reals^\Y_+$ \raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
	Since $f$ is defined over the simplex, and therefore $\sum_i p_i = 1$, we can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
	Moreover, since $f$ is polyhedral, it is finitely generated \jessie{Rockafellar 19.1.2} and can be written 
	\begin{align*}
	f(p) &= h(p) + \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex\jessie{?})\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

\item We now extend $f$ from the simplex to the nonnegative reals.
\begin{definition}
	Define the function $g : \reals_+^\Y \to \reals_+$ by extending $f$
	\begin{align*}
	g(x) = \min_{v\in\hat\V} \inprod{x}{v} - \delta(x \mid \reals_+^\Y)~,~
	\end{align*}
	taking $g(0) = 0$.
\end{definition}
%Extend to $g(x) = \min_{v\in\hat\V} \inprod{x}{v}$ for $x \in \reals^\Y_+$ and $g(x) = -\infty$ otherwise.  Def $g(0) = 0$.

@@

%\item Let $H_y^+ := \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Let $\H_{\hat \V} = \{H_v^+ \mid v\in\hat\V\}$, $\H_\Y = \{H_y^+ \mid y\in\Y\}$, $\H = \H_{\hat \V} \cup \H_\Y$.
\item
\begin{claim}
	$G = \cap \H$.
\end{claim}
\begin{proof}
	$(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \hat \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \hat \V$ and $x_y \geq 0$ for all $y$.
	Therefore $(x,c) \in H^+_v$ for all $v \in \hat \V$ and in $H^+_y$ for all $y \in \Y$, so $(x,c) \in \cap \H$.
\end{proof}
\end{itemize}

@@

\item \raf{This is the important general polyhedra result to cite/state.  Note: we'll have to work half-spaces of the form $H_{(w,\beta)}^+ = \{ z \in \reals^{n+1} \mid \inprod{z}{w} - \beta \geq 0\}$.}  Since $G$ is a full-dimensional polyhedron, with a unique H-rep by [cite].
  Hence, there is some unique $\H' \subseteq \H$ such that $G = \cap \H'$. \jessie{$\H \subseteq \H$ since $G = \cap \H' = \cap \H$; something about smallest cardinality here?}
  Moreover, each $H\in\H'$ defines a facet \raf{make formal}
  \begin{proof}
	\jessie{Incomplete}
	We know $G$ has a halfspace representation $G = G(H', \vec 0)$ uniquely determined by the facets of $G$ by \citep[Proposition 4.5]{gallier2008notes} \jessie{Another reference: Ziegler Proposition 2.15 analagous extension}.
	Take $\H' := \H_{H'}$; we claim that $\H' \subseteq \H$ and $G = \cap \H'$. \jessie{$\H' \subseteq \H$ seems to follow from Proposition 8.5 from \url{https://people.inf.ethz.ch/fukudak/lect/pclect/notes2014/PolyComp2014.pdf}}
  \end{proof}
\item 
\begin{claim}
	$\H_\Y \subseteq \H'$.
\end{claim}
\begin{proof}
	We argue $\H_\Y \subseteq \H'$; if there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H'$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H'$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
	The first cannot happen as we take $f$ to be defined (and finite) on $p = \delta_y \in \simplex$, and $f$ and $g$ match on the domain of $f$.
	Moreover, the second cannot be true by construction of $g$.	
\end{proof}
\item 
\begin{corollary}
	Then $\H' \setminus \H_\Y = \{H_v^+ : v\in\V\} =: \H_\V$ for some $\V \subseteq \hat \V$; That's the $\V$ we are looking for.  (And we've argued that it's uniquely determined by $f$.) \jessie{Why is $\V$ minimum cardinality?}
\end{corollary}
\item Define $F_v := H_v \cap G$ for any $v\in\reals^\Y_+$.
\item Claim: $F_v$ is a facet of $G$ for each $v\in\V$.  \raf{Unpacking ``defines a facet''}
\end{itemize}

Projections to the positive orthant
\begin{itemize}
\item Define the projection $\pi:\reals^\Y\times \reals \to \reals^\Y, (x,c) \mapsto x$.
\item Claim: For all $x\in\reals^\Y_+$, there exists $v\in\V$ such that $H_v$ supports $G$ at $(x,g(x))$.
\item Corollary: $\cup_{v\in\V} \pi(F_v) = \reals^\Y_+$. \jessie{Not sure how to show this, but intuitively, this is saying that the projection of the facets union to $\reals^\Y_+$.}
\item Note: $\theta(v) = \pi(F_v) \cap \simplex$. \raf{But haven't def $\theta$ yet}
\end{itemize}

>>>>

\section{``Book chapter'' approach}
\subsection*{Definitions and preliminaries}
\begin{definition}[Closed halfspace]
	For any $w\in\reals^\Y_+, b \in \reals$, let $H_{(w,b)}^+ := \{ x \in \reals^d \mid \inprod{x}{w} - b \geq 0\}$ be the closed halfspace defined by $(w,b)$, where $d = |\Y| +1$.
	Abusing notation, we sometimes write $H^+_v := H^+_{([v;-1], 0)}$ for a $v \in \reals^\Y_+$.\jessie{Is notation clear?}
\end{definition}
Additionally, we sometimes let $H_y^+ := H_{e_y}^+$. % \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\begin{definition}[Hyperplane]
	For any $(w,b)\in\reals^\Y_+ \times \reals$, let $H_{(w,b)} := \{ x \in \reals^d \mid \inprod{x}{w} - b = 0\}$ be the hyperplane generated by $(w,b)$.
	Again, we let $H_v := H_{([v;-1], 0)}$ for a $v \in \reals^\Y_+$. 
\end{definition}

\begin{definition}[Polyhedron - halfspace representation\jessie{Ziegler page 28}]
	An \emph{polyhedron} $P$ describes an intersection of a finite set of closed halfspaces $\H = \{H^+_{(w_i,b_i)}\}$ presented in the form $P = \cap \H$.
%	\begin{equation*}
%	P = P(A,z) = \{x \in \reals^d \mid Ax \leq z\} \qquad \textrm{ for some } A \in \reals^{m \times d}, z \in \reals^m~.~
%	\end{equation*}
\end{definition}
\jessie{Comment the above out, and define as a finite intersection of closed halfspaces}
\begin{definition}[Face, facet]
	Let $P \subseteq \reals^d$ be a convex polytope.
	A halfspace $H^+_{(w,b)}$ is \emph{valid} for $P$ if $P \subseteq H^+_{(w,b)}$.
	%A linear inequality $\inprod{w}{x} \geq b$ is \emph{valid} for $P$ if it is satisfied for all $x \in P$.
	A \emph{face} of the polytope $P$ is any set of the form
	\begin{equation*}
	F = P \cap H_{(w,b)}~,~
	\end{equation*}
	for any valid halfspace $H^+_{(w,b)}$.
	The dimension of a face is the dimension of its affine hull $\dim(F) := \dim(\affhull(F))$.
	A $(\dim(P) - 1)$-dimensional face is called a facet.
\end{definition}

\begin{theorem}\label{thm:polyhedron-uniquely-gen-facets}
	Given a $d$-dimensional polyhedron $P \subseteq \reals^d$, there is a unique finite set of halfspaces $\H = \{H^+_{(w_1, b_1)}, \ldots, H^+_{(w_k, b_k)}\}$ such that $P = \cap \H$, and for all sets of closed halfspaces $\H'$ such that $P = \cap \H'$, we have $\H \subseteq \H'$.
	Moreover, for each $i \in [k]$, letting $F_i := H_{(w_i, b_i)} \cap P$, the set of $\{F_1, \ldots, F_k\}$ is the set of facets of $P$.
\end{theorem}
\begin{proof}
	\jessie{TODO}
\end{proof}
\jessie{Need to find a good reference.  Cite a result for polytopes and transform the polyhedra into a polytope. Could state the result for class of polyhedra containing $G$.
	Add two facets and now it's a polytope.
}

@@

Set up notation and define $g,G$
\begin{itemize}\setlength{\itemsep}{0pt}
\item Let $f:\simplex\to\reals_+$ be a polyhedral concave function.
\begin{claim}
	We may write $f(p) = \min_{v\in\hat\V} \inprod{p}{v}$ for some finite set $\hat\V \subseteq \reals^\Y_+$ \raf{show you can do this using the $h$ and $\delta$ representation from Rockafellar -- argue that because $p\in\simplex$ we can write $\inprod{p}{b_i}-\beta_i = \inprod{p}{b_i-\beta_i\ones}$.}
\end{claim}
\begin{proof}
	Since $f$ is defined over the simplex, and therefore $\sum_i p_i = 1$, we can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
	Moreover, since $f$ is polyhedral, it is finitely generated \jessie{Rockafellar 19.1.2} and can be written 
	\begin{align*}
	f(p) &= h(p) - \delta(p \mid C)\\
		 &= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex\jessie{?})\\
		 &= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
	\end{align*}
\end{proof}

\item We now extend $f$ from the simplex to the nonnegative reals.
\begin{definition}
	Define the function $g : \reals_+^\Y \to \reals_+$ by extending $-f$
	\begin{align*}
	g(x) = \min_{v\in\hat\V} \inprod{x}{v} - \delta(x \mid \reals_+^\Y)~,~
	\end{align*}
	taking $g(0) = 0$.
\end{definition}
%Extend to $g(x) = \min_{v\in\hat\V} \inprod{x}{v}$ for $x \in \reals^\Y_+$ and $g(x) = -\infty$ otherwise.  Def $g(0) = 0$.

@@

%\item Let $H_y^+ := \{ (x,c) \in \reals^\Y \times \reals \mid x_y \geq 0\}$.
\item Let $\H_{\hat \V} = \{H_v^+ \mid v\in\hat\V\}$, $\H_\Y = \{H_y^+ \mid y\in\Y\}$, $\H = \H_{\hat \V} \cup \H_\Y$.
\item
\begin{claim}
	$G = \cap \H$.
\end{claim}
\begin{proof}
	$(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \hat \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \hat \V$ and $x_y \geq 0$ for all $y$.
	Again, this is if and only if $(x,c) \in H^+_v$ for all $v \in \hat \V$ and in $H^+_y$ for all $y \in \Y$, so $(x,c) \in \cap \H$.
\end{proof}
\end{itemize}

@@

\item 
\begin{claim}
	\raf{This is the important general polyhedra result to cite/state.  Note: we'll have to work half-spaces of the form $H_{(w,\beta)}^+ = \{ z \in \reals^{n+1} \mid \inprod{z}{w} - \beta \geq 0\}$.}  
	%Since $G$ is a full-dimensional polyhedron, with a unique H-rep by [cite].
	There is some unique $\H' \subseteq \H$ such that $G = \cap \H'$. \jessie{$\H \subseteq \H$ since $G = \cap \H' = \cap \H$; something about smallest cardinality here?}
	Moreover, for each $H_{(w_i, b_i)} \in\H'$, the face $F_{(w_i, b_i)} = H_{(w_i, b_i)} \cap G$ is a facet. %\raf{make formal}
\end{claim}
  \begin{proof}
  	Since $G$ is full-dimensional, this follows immediately from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}.
%	We know $G$ has a halfspace representation $G = G(H', \vec 0)$ uniquely determined by the facets of $G$ by \citep[Proposition 4.5]{gallier2008notes} \jessie{Another reference: Ziegler Proposition 2.15 analagous extension}.
%	Take $\H' := \H_{H'}$; we claim that $\H' \subseteq \H$ and $G = \cap \H'$. \jessie{$\H' \subseteq \H$ seems to follow from Proposition 8.5 from \url{https://people.inf.ethz.ch/fukudak/lect/pclect/notes2014/PolyComp2014.pdf}}
  \end{proof}
\item 
\begin{claim}
	$\H_\Y \subseteq \H'$.
\end{claim}
\begin{proof}
	We argue $\H_\Y \subseteq \H'$; if there was a $y \in \Y$ such that $H_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H'$, then we would either have some $c_1 > 0$ such that $\{(x,c) \mid x_y \geq c_1\} \in \H'$, or we have a point $x$ such that $x_y < 0$ but $g(x) > -\infty$.
	The first cannot happen as we take $f$ to be defined (and finite) on $p = \delta_y \in \simplex$, and $f$ and $g$ match on the domain of $f$.
	Moreover, the second cannot be true by construction of $g$.	
\end{proof}
\item 
\begin{corollary}
	Then $\H' \setminus \H_\Y = \{H_v^+ : v\in\V\} =: \H_\V$ for some $\V \subseteq \hat \V$; That's the $\V$ we are looking for.  (And we've argued that it's uniquely determined by $f$.) \jessie{Why is $\V$ minimum cardinality?}
\end{corollary}

\item Define $F_v := H^+_v \cap G$ for any $v\in\reals^\Y_+$.
\item 
\begin{claim}
	$F_v$ is a facet of $G$ for each $v\in\V$.  \raf{Unpacking ``defines a facet''}
\end{claim}
\begin{proof}
This follows immediately since $\H_\V \subseteq \H'$ and every $H^+ \in \H'$ defines a facet by Theorem~\ref{thm:polyhedron-uniquely-gen-facets}.
%	First, observe that by construction, $H_v = \{(x,c) \mid \inprod{v}{x} - c \geq 0\}$ is a valid inequality for $G$, and thus $F_v$ is a face of $G$ by definition.
%	To see that $F_v$ is a facet, we must show it is $\dim(G) - 1 = |\Y|$-dimensional.
%	This holds true if we can fit a $|\Y|$-dimensional ball in $F_v$.
%	If we could not fit such a ball in $F_v$, \jessie{contradict $\V$ minimum cardinality}
\end{proof}
\end{itemize}

Projections to the positive orthant
\begin{itemize}
\item Define the projection $\pi:\reals^\Y\times \reals \to \reals^\Y, (x,c) \mapsto x$.
\item 
\begin{claim}
	For all $x\in\reals^\Y_+$, there exists $v\in\V$ such that $H_v$ supports $G$ at $(x,g(x))$.
\end{claim}
\begin{proof}
We can write any face $F_v = \{x \in \reals^d \mid \inprod{x}{v} = g(x)\}$.
If there was no $v \in \V$ such that $x \in F_V$, then $g(x) = \min_{H_v \in \H_\Y} \inprod{v}{x}$.
As these halfspaces are vertical, we have $g(x) = -\infty$, which does not support $G$ at $(x, g(x))$ since $x \in \reals_+^\Y$, yielding a contradiction.
\jessie{Don't like this.}
%We claim $\cup_{v \in \V}\pi(F_v) = \reals_+^\Y$.
%If there was a $v \in \V$ such that $\pi(F_v) \not \subseteq \reals_+^\Y$, this would imply there is some $x \not \in \reals_+^\Y$ such that $\inprod{x}{v} = c < \infty = g(x)$, contradicting $F_v$ supporting $g$ at $x$.
%Thus, $\cup_{v \in \V} \pi(F_v) \subseteq \reals_+^\Y$.
%Now for the other direction of subset inclusion, observe that for all $x \in \reals^\Y_+$, we have $g(x) = f(x) = \min_{v \in \V}\inprod{v}{x}$, so there is a $v \in \V$ supporting $g$ at $x$, and therefore $x \in \pi(F_v)$ for some $v \in \V$, and is therefore in the union of projections.
%As we have shown both directions of set inclusion, we must have $\cup_{v \in \V}\pi(F_v) = \reals_+^\Y$.	
\end{proof}

\item 
\begin{corollary}
	$\cup_{v\in\V} \pi(F_v) = \reals^\Y_+$.
	% \jessie{Not sure how to show this, but intuitively, this is saying that the projection of the facets union to $\reals^\Y_+$.}
\end{corollary}
\item Note: $\theta(v) = \pi(F_v) \cap \simplex$. \raf{But haven't def $\theta$ yet}
\end{itemize}
